SU($n$) spin models near the Heisenberg critical point

### 2020-03-27 collective_sun/sun_perturbation.tex
added discussion of the algorithm to construct a basis for a multi-body excitation manifold

The PS manifold $\M_0$ is spanned by a basis of states
$\ket{m}=\ket{m_1,m_2,\cdots,m_n}$ with a definite occupation number
$m_\mu$ of the single-spin state $\mu$, and $\sum_\mu m_\mu=N$.  In
the case of $n=2$, for example, the PS manifold $\M_0$ is precisely
the Dicke manifold\cite{dicke1954coherence} of $N$ qubits, spanned by
Dicke states $\set{\ket{m_0,m_1}}$ for which $m_0+m_1=N$ and the
integer $m_0$ ($m_1$) indicates the total occupation number of qubit
state $\ket{0}$ ($\ket{1}$).  The dimension of the PS manifold $\M_0$
is determined by the number of ways to assign $N$ spins to $n$ states,
which is ${N+n-1 \choose n-1}\sim N^{n-1}$.  The energy of any PS
state $\ket\psi\in\M_0$ with respect to the interaction Hamiltonian
$H_0$ in \eqref{eq:H_0} is
\begin{align}
  E_0 = \sum_{\set{p,q}\in\C_N\p{2}} h_{pq}.
\end{align}

@@

The second order effective Hamiltonian $H_{\t{eff}}^{(2)}$ is not as
straightforward to simplify due to the presence of a projector
$\P_\Delta$ onto the manifold $\E_\Delta$ of states with excitation
energy $\Delta$.  For each term $X_{\m w}$ of $\hat\O$, this projector
essentially picks off the part of $X_{\m w}$ that is strictly
off-diagonal with respect to the ground- and excited-state manifolds
$\E_0$ and $\E_\Delta$.  To decompose $X_{\m w}$ into components that
generate states of definite excitation energy when acting on PS states
$\M_0$, we pick an arbitrary state $\ket\psi\in\M_0$ and expand (see
Appendix \ref{sec:eigenstates})
\begin{align}
  H_0 X_{\m w} \ket\psi
  = E_0 X_{\m w} \ket\psi + X_{\hat{\m h}\p{\m w}} \ket\psi,
  \label{eq:diagnosis}
\end{align}
where $\hat{\m h}$ is a linear map on the space of dimension-$M$
tensors, determined entirely by the couplings $h_{pq}$ in $H_0$.  The
image $\hat{\m h}\p{\m w}$ of $\m w$ under $\hat{\m h}$ is then itself
a dimension-$M$ tensor that defines the $M$-body operator
$X_{\hat{\m h}\p{\m w}}$.  To write out $\hat{\m h}\p{\m w}$
explicitly, we first define a matrix
$\m h\equiv\sum_{p,q\in\ZZ_N} h_{pq} \op{p}{q}$ of all couplings in
$H_0$ by enforcing that $h_{pq}=h_{qp}$ and $h_{pp}=0$.  We denote a
contraction of $\m h$ with the $a$-th index of $\m w$ by
$\m h_a\p{\m w}$, such that the $k$-th entry of $\m h_a\p{\m w}$ is
\begin{align}
  \m h_a\p{\m w}_k
  \equiv \sum_{\substack{p\in\ZZ_N\\p\notin k}}
  h_{pk_a} w_{k_1 k_2 \cdots k_{a-1} p k_{a+1} \cdots k_M}
  = \sum_{\substack{p\in\ZZ_N\\p\notin k}} h_{pk_a} w_{k_{a:p}},
\end{align}
where if $k=\p{k_1,k_2,\cdots,k_M}$, then $k_{a:p}$ is the same list
but with the $a$-th entry $k_a$ replaced by $p$.  We similarly define
a tensor $\m h_{\t{P}}\p{\m w}$ with entries
\begin{align}
  \m h_{\t{P}}\p{\m w}_k
  \equiv \sum_{\set{a,b}\in\C_M\p{2}} h_{k_ak_b} w_{k_{a\lra b}},
\end{align}
where $k_{a\lra b}$ is the same as $k$ but with the placement of $k_a$
and $k_b$ swapped.  Finally, we define a tensor $\m h_{\t{D}}\p{\m w}$
with entries
\begin{align}
  \m h_{\t{D}}\p{\m w}_k \equiv h^{\t{D}}_k w_k,
  &&
  h^{\t{D}}_k \equiv \sum_{p\in k} h_p
  - \sum_{\set{a,b}\in\C_M\p{2}} h_{k_ak_b},
  &&
  h_p \equiv \sum_{q\in\ZZ_N} h_{pq}.
\end{align}
These definitions allow us to expand
\begin{align}
  \hat{\m h}\p{\m w}
  \equiv \sum_{a\in\ZZ_M} \m h_a\p{\m w}
  + \m h_{\t{P}}\p{\m w} - \m h_{\t{D}}\p{\m w}.
\end{align}
Viewed an operator on the vector space of dimension-$M$ tensors, the
map $\hat{\m h}$ is real and symmetric under transposition, and is
therefore diagonalizable.  Diagonalizing this operator yields
eigenvalues $\Delta$ and eigenvectors $\m m_\Delta$ for which
$\hat{\m h}\p{\m m_\Delta}=\Delta\m m_\Delta$.  Finding eigenvalues
and eigenvectors of $\hat{\m h}$ nominally requires diagonalizing a
matrix of dimensions $\abs{\D_N\p{M}}\times\abs{\D_N\p{M}}$, but
spatial symmetries of $H_0$ and $X_{\m w}$ such as translational
invariance or isotropy can drastically reduce the complexity of this
eigenvalue problem.  We discuss these reductions in Appendix
\ref{sec:symmetries}.

@@

In order to find the effective Hamiltonian $H_{\t{eff}}^{(2)}$ induced
by the multi-body operator $\hat\O$ on the PS manifold $\M_0$ at
second order in perturbation theory, we constructed states with
definite excitation energy $\Delta$ with respect to $H_0$.  These
states are generated by the action of an $M$-body operator
$X_{\m w_\Delta}$ on a PS state $\ket\psi\in\M_0$.  We denote the
manifold of all states that can be generated by the action of some
$M$-body operator, but not any $r$-body operator with $r<M$, by
$\M_M$; we will refer to $\M_M$ as the $M$-body excitation manifold.
We then denote the a direct sum of all $\M_r$ for $r\le M$ by
$\ul{\M}_M$, i.e.~$\ul{\M}_M\equiv\bigoplus_{r\le M}\M_r$.  The
manifold $\ul{\M}_M$ consists of all states accessible by the action
of $r$-body operators with $r\le M$ on PS states in $\M_0$.

By construction, the $M$-body operator $X_{\m w}$ generally couples PS
states in $\M_0$ to asymmetric states in $\ul{\M}_M\setminus\M_0$.  If
the coupling between $\M_0$ and its complement is weak compared to the
spectral gap $\Delta_{\t{gap}}$ of $H_0$, then we can treat the effect
of $X_{\m w}$ perturbatively and derive an effective Hamiltonian
$H_{\t{eff}}$ on $\M_0$.  This ground-state perturbation theory
formally breaks down when the operator norm
$\norm{X_{\m w}}\ge\Delta_{\t{gap}}/2$, in which case the operator
$X_{\m w}$ first couples $\M_0$ to $\ul{\M}_M$, then $\ul{\M}_M$ to
$\ul{\M}_{2M}$, and so on.  Even if
$\norm{X_{\m w}}>\Delta_{\t{gap}}/2$, however, the energetic penalty
incurred by $H_0$ from breaking the permutational symmetry of an
initial state $\ket\psi\in\M_0$ suppresses the leakage of population
into manifolds $\M_r$ of increasing $r$.  One might therefore capture
the dynamics of an initial PS state $\ket\psi\in\M_0$ by projecting
$X_{\m w}$ onto $\ul{\M}_{r_{\t{max}}}$ for some $r_{\t{max}}\ge M$,
resulting in a theory that resembles first-order perturbation theory
on $\ul{\M}_{r_{\t{max}}}$.  In the perturbative regime
$\norm{X_{\m w}}<\Delta_{\t{gap}}/2$, such a theory captures all
virtual processes through order $2\floor{r_{\t{max}}/M}+1$ in a formal
perturbation theory on $\M_0$.  In the non-perturbative regime
$\norm{X_{\m w}}\ge\Delta_{\t{gap}}/2$, the error incurred by
truncating the Hilbert space of a spin system past
$\ul{\M}_{r_{\t{max}}}$ is not clear a priori.  Nonetheless,
simulations restricted to $\ul{\M}_{r_{\t{max}}}$ can be benchmarked
by examining the maximum populations within all $\M_r$ for
$r\le r_{\t{max}}$ throughout simulation.  As long as
\begin{enumerate*}
\item these
populations fall off with increasing $r$,
\item and the maximal population within large-$r$ manifolds $\M_r$ is
  small,
\end{enumerate*}
one should expect such simulations to faithfully capture (at least
qualitatively) the dynamical behavior of the spin system under
consideration.

In order to project an operator $X_{\m w}$ onto
$\ul{\M}_{r_{\t{max}}}$, we first need to construct a basis for
$\ul{\M}_{r_{\t{max}}}$, and then compute all matrix elements of
$X_{\m w}$ in the constructed basis.  Our strategy will be to first
identify a basis for $\M_0$, and then find, for each
$r=1,\cdots,r_{\t{max}}$, a suitable set of $r$-body operators
$\set{O}$ and dimension-$r$ tensors $\set{\m u}$ for which the states
$\ket{O,\m u,m}\propto O_{\m u}\ket{m}$ form an orthonormal basis for
$\M_r$.  Matrix elements of $X_{\m w}$ with respect to this basis for
$\ul{\M}_{r_{\t{max}}}$ then take the form
\begin{align}
  \bk{O,\m u,\ell|X_{\m w}|Q,\m v,m}
  = \f{\bk{\ell|O_{\m u}^\dag X_{\m w} Q_{\m v}|m}}
  {\sqrt{\bk{\ell|O_{\m u}^\dag O_{\m u}|\ell}
      \bk{m|Q_{\m v}^\dag Q_{\m v}|m}}}.
  \label{eq:matrix_element}
\end{align}
In order to compute these matrix elements, we simplify products of the
form $\P_0 O_{\m u}^\dag X_{\m w} Q_{\m v} \P_0$ and
$\P_0 O_{\m u}^\dag O_{\m u} \P_0$, with $\P_0$ a projector onto
$\M_0$, in Appendix \ref{sec:operator_product}.

A suitable basis for $\M_0$ is the set of PS states
$\ket{m}=\ket{m_1,m_2,\cdots,m_n}$ with a definite occupation number
$m_\mu$ of the single-spin state $\mu$.

@@

% For consistency with \eqref{eq:matrix_element}, for each state
% $\ket{m}$ in this basis for $\M_0$ we define
% $\ket{\1,1,m}\equiv\ket{m}$, with $\1$ the identity operator, and we
% define $\1_1\equiv\1$.

@@

  \label{eq:multi_body_op}

@@

The process of constructing and simplifying diagrams to compute
coefficients of the diagrammatic operator product expansion in
\eqref{eq:sym_prod_diagrams} is systematic enough to be converted into
a computer algorithm.  The last step necessary to automate the
evaluation of \eqref{eq:sym_prod_diagrams} is the expansion of the
multi-spin operators $\m X\p{\ell}$ in a basis for the PS manifold
$\M_0$.  Without loss of generality, we consider a single $M$-spin
operator $X$; our task is essentially to find coefficients of the
expansion
\begin{align}
  \P_0 X \P_0
  = \sum_{a,b\in\B_0} \bk{a|X|b} \op{a}{b},
\end{align}
where $\set{\ket{a}:a\in\B_0}$ is complete bases for $\M_0$, and the
choice of $M$ spins on which $X$ acts is arbitrary.  In a system of
$N$ spins with $n$ states each, a suitable basis for $\M_0$ can be
labeled by the occupation number $a_\mu$ of each single-spin state,
namely $a=\p{a_1,a_2,\cdots,a_n}$ with $\sum_\mu a_\mu=N$.  Written
out explicitly,
\begin{align}
  \ket{a} = \f1{\sqrt{\C\p{a}}}
  \sum_{\substack{\t{distinct}\\\t{permutations}\\\Pi~\t{of}~\tilde a}}
  \Pi \ket{\tilde a},
  &&
  \ket{\tilde a} \equiv \bigotimes_\mu \ket{\mu}^{\otimes a_\mu},
  &&
  \C\p{a} \equiv \f{\p{\sum_\mu a_\mu}!}{\prod_\mu a_\mu!},
\end{align}
where the multinomial coefficient $\C\p{a}$ counts the number of
distinct ways to permute the tensor factors of $\ket{\tilde a}$, and
enforces $\bk{a|a}=1$.  We denote the set of all valid assignments of
$N$ (identical) spins to $n$ (distinguishable) states by $\A_n\p{N}$,
such that the basis $\set{\ket{a}:a\in\A_n\p{N}}$ spans
permutationally symmetric manifold of $N$ spins.  Identifying this
basis allows us to expand
\begin{align}
  \bk{a|X|b}
  = \sum_{\substack{\alpha,\beta\in\A_n\p{M}\\\alpha\le a,\,\beta\le b}}
  \delta_{a-\alpha,b-\beta}
  \sqrt{\f{\C\p{\alpha}\C\p{a-\alpha}\C\p{\beta}\C\p{b-\beta}}
    {\C\p{a}\C\p{b}}}
  \bk{\alpha|X|\beta},
  \label{eq:multi_body_eval}
\end{align}
where the restriction $\alpha\le a$ and difference $a-\alpha$ are
evaluated element-wise, i.e.~$\alpha<a\implies \alpha_\mu\le a_\mu$
and $\p{a-\alpha}_\mu=a_\mu-\alpha_\mu$ for all $\mu$; and
$\delta_{cd}=1$ if $c=d$ and zero otherwise.  We sum over both
$\alpha$ and $\beta$ above merely to keep the expression symmetric
with respect to transposition; in practice, one can simply sum over
$\alpha\in\A_n\p{M}$ and set $\beta=b-a+\alpha$.  Note that, by slight
abuse of notation, the operator $X$ on the left of
\eqref{eq:multi_body_eval} acts on an arbitrary choice of $M$ spins
(out of $N$), whereas the operator $X$ on the right of
\eqref{eq:multi_body_eval} is simply an $M$-spin operator, with matrix
elements $\bk{\alpha|X|\beta}$ evaluated with respect to the $M$-spin
states $\ket\alpha,\ket\beta$.

>>>>

The PS manifold $\M_0$ is spanned by a basis of states
$\ket{m}=\ket{m_1,m_2,\cdots,m_n}$ with a definite occupation number
$m_\mu$ of the single-spin state $\mu$, and $\sum_\mu m_\mu=N$.  We
denote the set of all such assignments of $N$ (identical) spins to $n$
(distinct) states by $\A_n\p{N}$, such that our standard basis for
$\M_0$ is $\set{\ket{m}:m\in\A_n\p{N}}$.  The dimension of $\M_0$ is
then the number of such assignments, which is straightforward to
determine with a standard ``stars and bars'' calculation:
$\abs{\A_n\p{N}} = {N+n-1 \choose n-1} \sim N^{n-1}$.  In the case of
$n=2$, for example, the PS manifold $\M_0$ is precisely the Dicke
manifold\cite{dicke1954coherence} of $N$ qubits, spanned by
${N+2-1\choose2-1}=N+1$ Dicke states $\ket{m_0,m_1}$ for which
$m_0+m_1=N$ and the integer $m_0$ ($m_1$) indicates the total
occupation number of qubit state $\ket{0}$ ($\ket{1}$).  The energy of
any PS state $\ket\psi\in\M_0$ with respect to the interaction
Hamiltonian $H_0$ in \eqref{eq:H_0} is
\begin{align}
  E_0 = \sum_{\set{p,q}\in\C_N\p{2}} h_{pq}.
\end{align}

@@

The second order effective Hamiltonian $H_{\t{eff}}^{(2)}$ is not as
straightforward to simplify due to the presence of a projector
$\P_\Delta$ onto the manifold $\E_\Delta$ of states with excitation
energy $\Delta$.  For each term $X_{\m w}$ of $\hat\O$, this projector
essentially picks off the part of $X_{\m w}$ that is strictly
off-diagonal with respect to the ground- and excited-state manifolds
$\E_0$ and $\E_\Delta$.  To decompose $X_{\m w}$ into components that
generate states of definite excitation energy when acting on PS states
$\M_0$, we pick an arbitrary state $\ket\psi\in\M_0$ and expand (see
Appendix \ref{sec:eigenstates})
\begin{align}
  H_0 X_{\m w} \ket\psi
  = E_0 X_{\m w} \ket\psi + X_{\hat{\m h}\p{\m w}} \ket\psi,
  \label{eq:diagnosis}
\end{align}
where $\hat{\m h}$ is a linear map on the space of dimension-$M$
tensors, determined entirely by the couplings $h_{pq}$ in $H_0$.  The
image $\hat{\m h}\p{\m w}$ of $\m w$ under $\hat{\m h}$ is then itself
a dimension-$M$ tensor that defines the $M$-body operator
$X_{\hat{\m h}\p{\m w}}$.  To write out $\hat{\m h}\p{\m w}$
explicitly, we first define a matrix
$\m h\equiv\sum_{p,q\in\ZZ_N} h_{pq} \op{p}{q}$ of all couplings in
$H_0$ by enforcing that $h_{pq}=h_{qp}$ and $h_{pp}=0$.  We denote a
contraction of $\m h$ with the $a$-th index of $\m w$ by
$\m h_a\p{\m w}$, such that the $k$-th entry of $\m h_a\p{\m w}$ is
\begin{align}
  \m h_a\p{\m w}_k
  \equiv \sum_{\substack{p\in\ZZ_N\\p\notin k}}
  h_{pk_a} w_{k_1 k_2 \cdots k_{a-1} p k_{a+1} \cdots k_M}
  = \sum_{\substack{p\in\ZZ_N\\p\notin k}} h_{pk_a} w_{k_{a:p}},
\end{align}
where if $k=\p{k_1,k_2,\cdots,k_M}$, then $k_{a:p}$ is the same list
but with the $a$-th entry $k_a$ replaced by $p$.  We similarly define
a tensor $\m h_{\t{P}}\p{\m w}$ with entries
\begin{align}
  \m h_{\t{P}}\p{\m w}_k
  \equiv \sum_{\set{a,b}\in\C_M\p{2}} h_{k_ak_b} w_{k_{a\lra b}},
\end{align}
where $k_{a\lra b}$ is the same as $k$ but with the placement of $k_a$
and $k_b$ swapped.  Finally, we define a tensor $\m h_{\t{D}}\p{\m w}$
with entries
\begin{align}
  \m h_{\t{D}}\p{\m w}_k \equiv h^{\t{D}}_k w_k,
  &&
  h^{\t{D}}_k \equiv \sum_{p\in k} h_p
  - \sum_{\set{a,b}\in\C_M\p{2}} h_{k_ak_b},
  &&
  h_p \equiv \sum_{q\in\ZZ_N} h_{pq}.
\end{align}
These definitions allow us to expand
\begin{align}
  \hat{\m h}\p{\m w}
  \equiv \sum_{a\in\ZZ_M} \m h_a\p{\m w}
  + \m h_{\t{P}}\p{\m w} - \m h_{\t{D}}\p{\m w}.
  \label{eq:multi_body_op}
\end{align}
Viewed an operator on the vector space of dimension-$M$ tensors, the
map $\hat{\m h}$ is real and symmetric under transposition, and is
therefore diagonalizable.  Diagonalizing this operator yields
eigenvalues $\Delta$ and eigenvectors $\m u_\Delta$ for which
$\hat{\m h}\p{\m u_\Delta}=\Delta\m u_\Delta$.  Finding eigenvalues
and eigenvectors of $\hat{\m h}$ nominally requires diagonalizing a
matrix of dimensions $\abs{\D_N\p{M}}\times\abs{\D_N\p{M}}$, but
spatial symmetries of $H_0$ and $X_{\m w}$ such as translational
invariance or isotropy can drastically reduce the complexity of this
eigenvalue problem.  We discuss these reductions in Appendix
\ref{sec:symmetries}.

@@

In order to find the effective Hamiltonian $H_{\t{eff}}^{(2)}$ induced
by the multi-body operator $\hat\O$ on the PS manifold $\M_0$ at
second order in perturbation theory, we constructed states with
definite excitation energy $\Delta$ with respect to $H_0$.  These
states are generated by the action of an $M$-body operator
$X_{\m w_\Delta}$ on a PS state $\ket\psi\in\M_0$.  We denote the
manifold of all states that can be generated by the action of some
$M$-body operator, but not any $r$-body operator with $r<M$, by
$\M_M$; we will refer to $\M_M$ as the $M$-body excitation manifold.
We then denote the a direct sum of all $\M_r$ for $r\le M$ by
$\ul\M_M$, i.e.~$\ul\M_M\equiv\bigoplus_{r\le M}\M_r$.  The manifold
$\ul\M_M$ consists of all states accessible by the action of $r$-body
operators with $r\le M$ on PS states in $\M_0$.

By construction, the $M$-body operator $X_{\m w}$ generally couples PS
states in $\M_0$ to asymmetric states in $\ul\M_M\setminus\M_0$.  If
the coupling between $\M_0$ and its complement is weak compared to the
spectral gap $\Delta_{\t{gap}}$ of $H_0$, then we can treat the effect
of $X_{\m w}$ perturbatively and derive an effective Hamiltonian
$H_{\t{eff}}$ on $\M_0$.  This ground-state perturbation theory
formally breaks down when the operator norm
$\norm{X_{\m w}}\ge\Delta_{\t{gap}}/2$, in which case the operator
$X_{\m w}$ first couples $\M_0$ to $\ul\M_M$, then $\ul\M_M$ to
$\ul\M_{2M}$, and so on.  Even if
$\norm{X_{\m w}}>\Delta_{\t{gap}}/2$, however, the energetic penalty
incurred by $H_0$ from breaking the permutational symmetry of an
initial state $\ket\psi\in\M_0$ suppresses the leakage of population
into manifolds $\M_r$ of increasing $r$.  One might therefore capture
the dynamics of an initial PS state $\ket\psi\in\M_0$ by projecting
$X_{\m w}$ onto $\ul\M_{r_{\t{max}}}$ for some $r_{\t{max}}\ge M$,
resulting in a theory that resembles first-order perturbation theory
on $\ul\M_{r_{\t{max}}}$.  In the perturbative regime
$\norm{X_{\m w}}<\Delta_{\t{gap}}/2$, such a theory captures all
virtual processes through order $2\floor{r_{\t{max}}/M}+1$ in a formal
perturbation theory on $\M_0$.  In the non-perturbative regime
$\norm{X_{\m w}}\ge\Delta_{\t{gap}}/2$, the error incurred by
truncating the Hilbert space of a spin system past
$\ul\M_{r_{\t{max}}}$ is not clear a priori.  Nonetheless, simulations
restricted to $\ul\M_{r_{\t{max}}}$ can be benchmarked by examining
the maximum populations within all $\M_r$ for $r\le r_{\t{max}}$
throughout simulation.  As long as
\begin{enumerate*}
\item these populations fall off with increasing $r$,
\item and the maximal population within large-$r$ manifolds $\M_r$ is
  small,
\end{enumerate*}
one should expect such simulations to faithfully capture (at least
qualitatively) the dynamical behavior of the spin system under
consideration.

In order to project an operator $X_{\m w}$ onto $\ul\M_{r_{\t{max}}}$,
we first need to construct a basis for $\ul\M_{r_{\t{max}}}$, and then
compute all matrix elements of $X_{\m w}$ in this basis.  Our strategy
will be to first identify a basis for $\M_0$, and then find, for each
$r=1,\cdots,r_{\t{max}}$, a suitable set of $r$-body operators
$\set{O}$ and dimension-$r$ tensors $\set{\m u}$ for which the states
$\ket{O,\m u,m}\propto O_{\m u}\ket{m}$ form an orthonormal basis for
$\M_r$.  Matrix elements of $X_{\m w}$ with respect to this basis for
$\ul\M_{r_{\t{max}}}$ then take the form
\begin{align}
  \bk{O,\m u,\ell|X_{\m w}|Q,\m v,m}
  = \f{\bk{\ell|O_{\m u}^\dag X_{\m w} Q_{\m v}|m}}
  {\sqrt{\bk{\ell|O_{\m u}^\dag O_{\m u}|\ell}
      \bk{m|Q_{\m v}^\dag Q_{\m v}|m}}}.
  \label{eq:matrix_element}
\end{align}
In order to compute these matrix elements, we simplify products of the
form $\P_0 O_{\m u}^\dag X_{\m w} Q_{\m v} \P_0$ and
$\P_0 O_{\m u}^\dag O_{\m u} \P_0$, with $\P_0$ a projector onto
$\M_0$, in Appendix \ref{sec:operator_product}.

Denoting all assignments of $N$ (identical) spins to $n$ (distinct)
states by $\A_n\p{N}$, a suitable basis for $\M_0$ is the set of PS
states $\set{\ket{m}:m\in\A_n\p{N}}$, where each
$m=\p{m_1,m_2,\cdots,m_n}$ specifies the occupation number $m_\mu$ of
the single-spin state $\mu$.  For consistency with the expression of
matrix elements in \eqref{eq:matrix_element}, for each
$\ket{m}\in\M_0$ we define $\ket{\1,1,m}\equiv\ket{m}$, with $\1$ the
identity operator, and we define $\1_1\equiv\1$.  We now need to
determine the set of $r$-body operators $\set{O}$ and dimension-$r$
tensors $\set{\m u}$ that define our basis for $\M_r$ with
$r=1,2,\cdots,r_{\t{max}}$.  As an initial guess for $\set{\m u}$, we
can choose the set $\set{\m u_\Delta}_{\m h,r}$ of dimension-$r$
tensors $\m u_\Delta$ that generate eigenstates of $H_0$, i.e.~for
which $\hat{\m h}\p{\m u_\Delta} = \Delta \m u_\Delta$, with
$\hat{\m h}$ defined in \eqref{eq:multi_body_op}.  The set of $r$-body
operators to choose is less obvious, but can be determined by an
exhaustive search over all $r$-body operators relevant to a particular
scenario.  Specifically, the purpose of projecting operators onto
$\col\M_{r_{\t{max}}}$ is to simulate the dynamics induced by some
operator $\hat\O$ of the form in \eqref{eq:pert}, constructed from a
set $\O=\set{\p{X,\m w}}$ containing multi-spin operators $X$.  In
order to find a basis for the {\it relevant} states in $\M_r$, one can
simply consider the set $\set{O}_{\O,r}$ all $r$-spin operators that
can be constructed from products of the multi-spin operators $X$ in
$\O$.

In general, the states $\ket{O,\m u_\Delta,m}$ for all
$\p{O,\m u_\Delta,m}\in\set{O}_{\O,r}\times\set{\m u_\Delta}_{\m h,r}
\times\A_n\p{N}$ will form an overcomplete basis for the relevant
states in $\col\M_r$.  Our last step is therefore to reduce this set
of states to a minimal basis $\B_r$ for $\M_r$.  To this end, we
construct each of $\B_1,\B_2,\cdots,\B_{r_{\t{max}}}$ in sequence,
such that when constructing $\B_r$ we already have a minimal basis
$\col\B_{r-1}\equiv\bigcup_{s<r-1}\B_s$ for $\col\M_{r-1}$.  To
construct $\B_r$, we then consider each
$\p{O,\m u_\Delta,m}\in\set{O}_{\O,r}\times\set{\m u_\Delta}_{\m h,r}
\times\A_n\p{N}$ one at a time.  If $\ket{O,\m u_\Delta,m}$ is
orthogonal to all states in $\col\B_{r-1}$, as well as all other
states we have added to $\B_r$ thus far, then we include it in $\B_r$.
Otherwise, we discard $\ket{O,\m u_\Delta,m}$.  There are two
considerations to keep in mind for this construction: first, one
should always throw out null states $\ket{O,\m u_\Delta,m}=0$, which
can be identified by a vanishing self-overlap,
$\bk{O,\m u_\Delta,m|O,\m u_\Delta,m}=0$.  Second, computing overlaps
to check orthogonality can be computationally intensive.  When
considering a candidate state $\ket{O,\m u_\Delta,m}$ to add to
$\B_r$, one can therefore restrict orthogonality checks to other
states $\ket{Q,\m v_\epsilon,\ell}$ with $\epsilon=\Delta$; if the
excitation energies $\epsilon\ne\Delta$, then these states are
guaranteed to be orthogonal.  Any two states with different occupation
numbers of single-spin states $\mu\in\ZZ_n$ are similarly guaranteed
to be orthogonal.

@@

@@

@@

The process of constructing and simplifying diagrams to compute
coefficients of the diagrammatic operator product expansion in
\eqref{eq:sym_prod_diagrams} is systematic enough to be converted into
a computer algorithm.  The last step necessary to automate the
evaluation of \eqref{eq:sym_prod_diagrams} is the expansion of the
multi-spin operators $\m X\p{\ell}$ in a basis for the PS manifold
$\M_0$.  Without loss of generality, we consider a single $M$-spin
operator $X$.  Our task is essentially to find coefficients of the
expansion
\begin{align}
  \P_0 X \P_0
  = \sum_{a,b\in\A_n\p{N}} \bk{a|X|b} \op{a}{b},
\end{align}
where $\A_n\p{N}$ is the set of all ways to assign $N$ (identical)
spins to $n$ (distinct) states, such that for any $a\in\A_n\p{N}$, the
state $\ket{a}=\ket{a_1,a_2,\cdots,a_n}$ is labeled by the occupation
number $a_\mu$ of state $\mu$, with $\sum_\mu a_\mu = N$.  Written out
explicitly,
\begin{align}
  \ket{a} = \f1{\sqrt{\C\p{a}}}
  \sum_{\substack{\t{distinct}\\\t{permutations}\\\Pi~\t{of}~\tilde a}}
  \Pi \ket{\tilde a},
  &&
  \ket{\tilde a} \equiv \bigotimes_\mu \ket{\mu}^{\otimes a_\mu},
  &&
  \C\p{a} \equiv \f{\p{\sum_\mu a_\mu}!}{\prod_\mu a_\mu!},
\end{align}
where the multinomial coefficient $\C\p{a}$ counts the number of
distinct ways to permute the tensor factors of $\ket{\tilde a}$, and
enforces $\bk{a|a}=1$.  Using these states, we can expand
\begin{align}
  \bk{a|X|b}
  = \sum_{\substack{\alpha,\beta\in\A_n\p{M}\\\alpha\le a,\,\beta\le b}}
  \delta_{a-\alpha,b-\beta}
  \sqrt{\f{\C\p{\alpha}\C\p{a-\alpha}\C\p{\beta}\C\p{b-\beta}}
    {\C\p{a}\C\p{b}}}
  \bk{\alpha|X|\beta},
  \label{eq:multi_body_eval}
\end{align}
where the restriction $\alpha\le a$ and difference $a-\alpha$ are
evaluated element-wise, i.e.~$\alpha<a\implies \alpha_\mu\le a_\mu$
and $\p{a-\alpha}_\mu=a_\mu-\alpha_\mu$ for all $\mu$; and
$\delta_{cd}=1$ if $c=d$ and zero otherwise.  We sum over both
$\alpha$ and $\beta$ above merely to keep the expression symmetric
with respect to transposition; in practice, one can simply sum over
$\alpha\in\A_n\p{M}$ and set $\beta=b-a+\alpha$.  Note that, by slight
abuse of notation, the operator $X$ on the left of
\eqref{eq:multi_body_eval} acts on an arbitrary choice of $M$ spins
(out of $N$), whereas the operator $X$ on the right of
\eqref{eq:multi_body_eval} is simply an $M$-spin operator, with matrix
elements $\bk{\alpha|X|\beta}$ evaluated with respect to the $M$-spin
states $\ket\alpha,\ket\beta$.
